Cognitive Context: Impact on the Perception of a Robot

### 2015-10-29 main.tex
APA: {subject->participant}

Measuring anthropomorphism raises
interesting methodological challenges, if only because people's perception of
robots change over the duration of their interaction with the
robots~\citep{lemaignan2014dynamics}. \citet{takayama_perspectives_2012}
discusses the effects of these changes and differentiates between what she calls
an \emph{in-the-moment} and a \emph{reflective} perspective on agency: an
\emph{in-the-moment} perspective would refer to one's most immediate
response/sense in a given situation whereas a \emph{reflective} perspective would
indicate one's reaction/sense of a situation based on a more in-depth
consideration and contemplation.  In the first phase of interaction with a
robot, people might respond ``mindlessly'' instead of responding
consciously~\citep{nass_machines_2000}, and only after what is called the
\emph{familiarization} period, one might respond in a more reflective manner
instead of an in-the-moment reaction: \citet{reeves_media_1996} showed for
instance that subjects interacting with technologies in ways close to how they
would behave towards people, would still deny any human-like interaction in a post-hoc
questionnaire, illustrating this reflective perspective.

@@

We explore here a different approach to measure anthropomorphism, based on
eye-tracking.  By looking at the eyes fixation on the head of the robot, we have
been able to demonstrate significantly different patterns in two experimental
conditions subtly different in terms of human-likeness. This eye-tracking based
metric is behavioural and hence does not suffer post-hoc reconstruction by the
subjects, possibly providing a better, ``noise-less'', access to the human
attitude towards the machine, and, as such, a better proxy to the cognitive and
affective bonds that the human establishes with the robot.

@@

    \caption{\small The study follows a 2$\times$2 design: the \emph{robot} vs
        \emph{human} condition is within subject, while the \emph{shallow
        cognitive context} \vs \emph{deep cognitive context} is between subject.
        Two stimuli (\emph{Picking} task, {\bf A}, and \emph{Pointing} task,
        {\bf B}) were shown to the participants, introduced by brief verbal
        commands, reproduced in the table. Figure~\ref{fig:stimuli} shows
        pictures of the four stimuli.}

@@

We ran the study with 56 subjects (mean age: 20.9, $SD=3.9$, 31 females and
25 males).  The participants were recruited amongst students of our university.
To avoid students with a high familiarity with robots and artificial intelligent
systems, Computer Science and Electronics curriculum were excluded. This was
\textit{a posteriori} comforted by the questionnaire responses: no participants
reported itself as being very familiar with robots, 8 reported some familiarity
and 7 out of 56 owned a domestic robot (vacuum cleaner).

As stated previously, the study follows a 2$\times$2 design
(table~\ref{table:design}): the \emph{deep} \vs \emph{shallow cognitive context}
condition was between-subject (30 subjects performed the \emph{deep cognitive}
tasks while 26 were in the \emph{shallow cognitive} condition); the \emph{robot}
\vs \emph{human} condition was within-subject (the order of stimuli presentation
was counter-balanced over the 56 participants). The study lasted in average 15
minutes per participant.

@@

Running an ANOVA against the change of anthropomorphic attributions
(\deltaant) over the duration of the experiment in the two cognitive
conditions, shows that \deltaant is greater for tasks with a deeper cognitive
context than that for the shallow cognitive tasks ($F[1,54] = 6.54, p < .05$,
figure~\ref{h4}). In other words, after being exposed to short videos of a robot
performing simple tasks, subjects primed with a deeper cognitive context report
a significantly higher tendency to anthropomorphize robots \emph{in general} (as
the questionnaire on anthropomorphic attributions explicitly referred to
\emph{robots in general}).

@@

The second contribution is a methodological one: we present in this article
a novel technique to assess anthropomorphic projections that relies on a
biometric measurement (eye-tracking) to compare gaze fixation durations on
the face of the robot. We cross-validate this new metric with existing,
established, questionnaires. Compared to current techniques (post-hoc
annotations of videos and questionnaires), this new approach is objective,
is less impacted by experimental biases (like the \emph{observer effect},
where the subject's behaviour is impacted by the fact he/she knows that
he/she is observed) and take place during the interaction itself
(\emph{in-the-moment} measurement).

@@

Besides, one may point that, in real-world experiments, the robot does not
always remain in the field of view of the subject, and, at the very least, the
body of the robot may be partially occluded at time.  This is certainly true,
and while new experiments would be welcome to validate it, we believe that our
method remains effective in these situations. First it relies on the
\emph{ratio} of the gaze fixations on the head \vs gaze fixations on other parts
of the body, thus compensating for out-of-sight robots; second, our method
\emph{does not} provide an \emph{absolute} measurement of the level
anthropomorphic perception, but rather a \emph{relative} measurement: within a
group of subject, it enables to classify subjects in term of their level of
projected anthropomorphism onto the robot.  This means that, while this method
can not be used to assess the perception of robots between different interaction
situations (\ie different experiments), it can be employed within a group a
subjects taking part to the same experiment.

>>>>

Measuring anthropomorphism raises
interesting methodological challenges, if only because people's perception of
robots change over the duration of their interaction with the
robots~\citep{lemaignan2014dynamics}. \citet{takayama_perspectives_2012}
discusses the effects of these changes and differentiates between what she calls
an \emph{in-the-moment} and a \emph{reflective} perspective on agency: an
\emph{in-the-moment} perspective would refer to one's most immediate
response/sense in a given situation whereas a \emph{reflective} perspective would
indicate one's reaction/sense of a situation based on a more in-depth
consideration and contemplation.  In the first phase of interaction with a
robot, people might respond ``mindlessly'' instead of responding
consciously~\citep{nass_machines_2000}, and only after what is called the
\emph{familiarization} period, one might respond in a more reflective manner
instead of an in-the-moment reaction: \citet{reeves_media_1996} showed for
instance that participants interacting with technologies in ways close to how they
would behave towards people, would still deny any human-like interaction in a post-hoc
questionnaire, illustrating this reflective perspective.

@@

We explore here a different approach to measure anthropomorphism, based on
eye-tracking.  By looking at the eyes fixation on the head of the robot, we have
been able to demonstrate significantly different patterns in two experimental
conditions subtly different in terms of human-likeness. This eye-tracking based
metric is behavioural and hence does not suffer post-hoc reconstruction by the
participants, possibly providing a better, ``noise-less'', access to the human
attitude towards the machine, and, as such, a better proxy to the cognitive and
affective bonds that the human establishes with the robot.

@@

    \caption{\small The study follows a 2$\times$2 design: the \emph{robot} vs
        \emph{human} condition is within subject, while the \emph{shallow
        cognitive context} \vs \emph{deep cognitive context} is between subjects.
        Two stimuli (\emph{Picking} task, {\bf A}, and \emph{Pointing} task,
        {\bf B}) were shown to the participants, introduced by brief verbal
        commands, reproduced in the table. Figure~\ref{fig:stimuli} shows
        pictures of the four stimuli.}

@@

We ran the study with 56 participants (mean age: 20.9, $SD=3.9$, 31 females and
25 males).  The participants were recruited amongst students of our university.
To avoid students with a high familiarity with robots and artificial intelligent
systems, Computer Science and Electronics curriculum were excluded. This was
\textit{a posteriori} comforted by the questionnaire responses: no participants
reported itself as being very familiar with robots, 8 reported some familiarity
and 7 out of 56 owned a domestic robot (vacuum cleaner).

As stated previously, the study follows a 2$\times$2 design
(table~\ref{table:design}): the \emph{deep} \vs \emph{shallow cognitive context}
condition was between-subjects (30 participants performed the \emph{deep cognitive}
tasks while 26 were in the \emph{shallow cognitive} condition); the \emph{robot}
\vs \emph{human} condition was within-subject (the order of stimuli presentation
was counter-balanced over the 56 participants). The study lasted in average 15
minutes per participant.

@@

Running an ANOVA against the change of anthropomorphic attributions
(\deltaant) over the duration of the experiment in the two cognitive
conditions, shows that \deltaant is greater for tasks with a deeper cognitive
context than that for the shallow cognitive tasks ($F[1,54] = 6.54, p < .05$,
figure~\ref{h4}). In other words, after being exposed to short videos of a robot
performing simple tasks, participants primed with a deeper cognitive context report
a significantly higher tendency to anthropomorphize robots \emph{in general} (as
the questionnaire on anthropomorphic attributions explicitly referred to
\emph{robots in general}).

@@

The second contribution is a methodological one: we present in this article
a novel technique to assess anthropomorphic projections that relies on a
biometric measurement (eye-tracking) to compare gaze fixation durations on
the face of the robot. We cross-validate this new metric with existing,
established, questionnaires. Compared to current techniques (post-hoc
annotations of videos and questionnaires), this new approach is objective,
is less impacted by experimental biases (like the \emph{observer effect},
where the behaviour of the participant is impacted by the fact he/she knows that
he/she is observed) and take place during the interaction itself
(\emph{in-the-moment} measurement).

@@

Besides, one may point that, in real-world experiments, the robot does not
always remain in the field of view of the participant, and, at the very least, the
body of the robot may be partially occluded at time.  This is certainly true,
and while new experiments would be welcome to validate it, we believe that our
method remains effective in these situations. First it relies on the
\emph{ratio} of the gaze fixations on the head \vs gaze fixations on other parts
of the body, thus compensating for out-of-sight robots; second, our method
\emph{does not} provide an \emph{absolute} measurement of the level
anthropomorphic perception, but rather a \emph{relative} measurement: within a
group of participants, it enables to classify participants in term of their level of
projected anthropomorphism onto the robot.  This means that, while this method
can not be used to assess the perception of robots between different interaction
situations (\ie different experiments), it can be employed within a group a
participants taking part to the same experiment.
